---
title: "Task2_543X6"
output: html_document
date: '2022-07-17'
---

```{r, setup, include=FALSE}
knitr::opts_chunk$set(root.dir="/Users/Mario/Post-doc/Multi100 Team/code and data/raw data/mmp_lamp",echo = F)
```

## R Markdown

We want to estimate how education level affects wage gains at migration, controlling for Country fixed effects. The dependent variable is the wage gains at migration:
```{r}
rm(list=ls())
wd <- "/Users/Mario/Post-doc/Multi100 Team/code and data/raw data/mmp_lamp" # change working directory
setwd(wd)
library(readxl)
library(ggplot2)
library(dplyr)
library(plm)
library(lfe)
library(lmtest)
library(car)
mData <- read_excel("wage_gain_table.xlsx") # load data
```

## Create variable for schooling and dummies to control for Countries
```{r}

# Create dummies for schooling

vSchool <- mData$edyrs
vNS <- as.numeric(vSchool<9) # no school
vSHS <- as.numeric(vSchool==9|vSchool==10|vSchool==11) # some high school but no degree
vHSD <- as.numeric(vSchool==12) # high school degree
vSC <- as.numeric(vSchool==13|vSchool==14|vSchool==15) # some college but no bachelor
vGRAD <- as.numeric(vSchool>=16) # bachelor or more

# Create dummies for Countries

vCOL <- as.numeric(mData$country=="COL")
vDOM <- as.numeric(mData$country=="DOM") 
vECU <- as.numeric(mData$country=="ECU")
vGTM <- as.numeric(mData$country=="GTM")
vHTI <- as.numeric(mData$country=="HTI")
vMEX <- as.numeric(mData$country=="MEX")
vNIC <- as.numeric(mData$country=="NIC")
vPER <- as.numeric(mData$country=="PER")
vSLV <- as.numeric(mData$country=="SLV")

```

## Create dependent variable: difference between US adjusted wage and home adjusted wage

```{r}

vWage <- abs(mData$lastUsWageAdjusted-mData$lastHomeWageAdjusted)

```

## I regress schooling on wage gains, including dummies to control for the effect of Country

```{r}
lEst <- lm(vWage~1+vNS+vSHS+vHSD+vSC
		       +vCOL+vDOM+vGTM+vHTI+vMEX+vNIC+vSLV)
summary(lEst)

```

Our aim is to test wether the wage gain at migration decreases as the number of years of schooling of the migrants increase.
The analysis shows that people who did not attend school (or went to school for less than 9 years in total) gain more at migration in term of wage increase. This is confirmed by the fact that the only significant variables are "no school" and "some high school".In particular, we can also note that the magnitude and the significance of the coefficients increase as the years of school decrease. To deal with Country fixed effects, we remove Peru as it is an outlier (it represents just a few dirty observations) and we create dummy variable for the other Countries. Since the Countries have all the same weights, Mexico is the only significant control variable as its sample size is much grater than that of the other Countries.

## I perform a joint Hypothesis testing, using the F-Statistic, to test wether the coefficients related to some high school but no degree (vSHS), high school degree (vHSD) and some college (vSC) are zero

```{r}

Ftest <- linearHypothesis(lEst,c("vSHS=0","vHSD=0","vSC=0"))
Ftest

```

The value of the F-statistic for the joint hypothesis test is approximately equal to 0.4029. Its p-value is approximately equal to 0.759>0.05. This means that we cannot reject the null hypothesis that the coefficients vSHS, vHSD and vSC are zero. Therefore, the final model is the following (controlling for Countries)


```{r}

lEst1 <- lm(vWage~1+vNS+vCOL+vDOM+vGTM+vHTI+vMEX+vNIC+vSLV)
summary(lEst1)

```