---
title: "Multi100-Analysis"
output: html_notebook
---
Re-Analysis of "Hurst_EvoHumanBehavior_2017_yypJ" as part of the Multi100-Project (https://osf.io/7snkz/)

$\rightarrow$ Raphael Merz, Ruhr-University Bochum (Raphael.Merz@rub.de), Analyst ID: PH39A

### set working-directory and load packages
```{r}
# setwd("~/Downloads/")

library(dplyr)
library(nortest)
library(car)
library(ggpubr)
library(lmtest)
```

### load data
```{r}
data <- read.csv("data.csv")
```

# First Look at the Data
```{r}
plot(data$DSM5_Total, data$MiniK_Total, pch = 16, col = "blue2")
```

# Analysis
## Preleminary Tests
### Liniarity

```{r}
plot(jitter(data$DSM5_Total) ~ jitter(data$MiniK_Total),xlab = "DSM-5-Scores", ylab = "Mini-K-Scores", las = 1, pch = 16, col = "blue2")
title("Life History and Psycholpathology")
lines(lowess(data$DSM5_Total ~ data$MiniK_Total),
      col = "red3", lwd = 3)
```

### Normality of the Mini-K
```{r}
plot(density(data$MiniK_Total, na.rm = TRUE))

curve(dnorm(x, mean(data$MiniK_Total), sd(data$MiniK_Total)), add = TRUE, lty = 2)

ggqqplot(data$MiniK_Total, ylab = "Mini-K-Scores")

shapiro.test(data$MiniK_Total)
```
$\rightarrow$ Normality is not given for the Mini-K-Questionnaire.

### Normality of the Mini-K
```{r}
plot(density(data$DSM5_Total, na.rm = TRUE))
curve(dnorm(x, mean(data$DSM5_Total), sd(data$DSM5_Total)), add = TRUE, lty = 2)

ggqqplot(data$DSM5_Total, ylab = "Mini-K-Scores")

shapiro.test(data$DSM5_Total)
```
$\rightarrow$ Normality is not given for the DSM-5-Questionnaire.

### Homoscedasticity
```{r}
leveneTest(dv~gr, data=rbind(data.frame(dv=data$DSM5_Total, gr='gr1'), data.frame(dv=data$MiniK_Total, gr='gr2')), center='mean')
# taken from: https://stats.stackexchange.com/questions/460348/levenes-test-in-two-different-samples.
```

$\rightarrow$ Homoscedasticity was not given either. Thus, a *Spearman Correlation* will be computed!

## Testing the Correlation
```{r}
cor.test(data$DSM5_Total, data$MiniK_Total, method = "spearman", alternative = "less")
```

# Conclusion
There is a significant negative correlation between life strategy and psychopathology within this study’s sample, *r*(136) = -.59, *p* < .001.