---
title: "Multi100_Anderson_AmEcoJourn_2011"
author: "David Wisniewski"
date: '`r Sys.Date()`'
output:
  html_document:
    theme: united
    highlight: tango
    code_folding: hide
    toc: true
    toc_float: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "../") # set the root directory

rm(list=ls(all=TRUE)) # clear work space
set.seed(42) # set seed for reproducible results
options(warn=-1) # do not print warnings to the html file

library(haven) # for .dta import
library(kableExtra) # for printing tables
library(ltm) # to compute cronbachs alpha
library(car) # used for the qq plots
library(MASS) # used for the GLMM

```


```{r data_import}

vil <- read_dta('Data/village.dta')
hh <- read_dta('Data/household.dta')
```
 
# The claim

In this study, the author assesses effects of dominant village caste on household level outcomes in two poor regions of rural India (Uttar Pradesh, Bihar). They compare household income for lower caste households in two types of villages: high caste dominated and low caste dominated. 

The claim to be tested here is the same as for Task 1: 

**"The analysis compares outcomes of lower caste (BAC, OBC, SC) households residing in both types of villages and finds substantially higher income for low-caste households residing in villages dominated by BACs." (p.240)**

Additional instruction for Task 2 are:

**Your analysis should produce a single, main result in terms of statistical families of z-, t-, F-, or χ² tests (or their alternative or non-parametric versions).You should not use District controls, Crop controls, Distance controls, Groundwater controls, Public goods controls Do not include Irrigation, credit, and tenancy variables, and their interaction term versions Do not use instruments for water buyer status.**

# Data and Operationalization

The data used is a World Bank survey from 1997-1998, from rural India, in which households from Uttar Pradesh and Bihar were included. A codebook of the variables included in this analysis is found below (Glossary). 

In order to test the claim, we first need to operationalize three key variables: lower caste households, village type, and income. Lower caste households will be those from the BAC, OBC, and SC castes, and only those will be included in the analysis. Village type will be operationalized through economic power, if the majority of the land is owned by high castes, the village is "high caste dominated". If the majority of the land is owned by low castes, the village is "low caste dominated". Income will be operationalized as the total household income, since the claim does not specify the source or type of income to be tested. 

# Analysis

## Analysis 1: Village level

The key analysis compares high vs low caste dominance villages, and how this affects household income. As a first step, we need to establish that these two village types only differ in caste dominance, and are equal in all other relevant village-level measures. The additional instructions seem targeted at the household level analysis below, which is why the village level analyis is reported here again. It is identical to the version from Task 1. 

The relevant variables included here were: irrigated land area, non-irrigated land area, landless households, electrified households, number of households, access to: hand pump, well, paved road, almost no land suffers from: flood, alkalinity, water logging, soil erosion, main crop is: paddy, wheat, price of: paddy, wheat, distance to: bus, telephone, police stations, pds shop, bank, primary school, middle school, secondary school, PHC, hospital, area in hectares, access to natural water. 

Any of these variables might differ between high and low caste dominated villages, e.g. showing that high caste dominated villages have better access to public goods. If this were the case, interpreting differences between both types of villages would become difficult. In order to test whether there were any difference in these variables, each of the listed variables was assessed separately. First, we computed a Shapiro Wilks test to assess if it was normally distributed, and results indicated that nop variable was normally distributed. In a next step, we performed a two-sample Mann Whitney U test for each variable, in order to test whether there were significant differences between both village types. 

``` {r village_comparison}
normal = NULL
ps = NULL
diffs = NULL
for (i in c(1:3,5,6,8:30)) { # for each village level variable
  res = shapiro.test(vil[[i]]) # test if variable is normally distributed
  
  if (res$p.value<0.05) { # if it is not, compute Mann Whitney U test
    mod = wilcox.test(vil[[i]] ~ vil[[7]], conf.int=T)
    diff = mod$estimate
    normal = rbind(normal,'no')
  } else { # if it is normally distributed, compute Welch t-test
    mod = t.test(vil[[i]] ~ vil[[7]], var.equal=F)
    diff = mod$estimate[1] - mod$estimate[2]
    normal = rbind(normal,'yes')
  }
  ps = rbind(ps,mod$p.value)
  diffs = rbind(diffs, diff)
}

df = data.frame(variable = names(vil)[c(-4,-7)],
                normal.distribution = normal,
                difference.in.location = as.numeric(diffs),
                pval.uncorrected = ps,
                pval.fdr = p.adjust(ps, method = 'fdr', n = length(ps)))

kbl(df,booktabs = T,escape =F, digits=c(0,0,6,2,2)) %>%
      kable_classic(full_width = F, html_font = "Cambria")

```

### Results 

The table shows the difference in location between both village types, the uncorrected p-values and FDR corrected p-values (for the total number of tests) of the Mann Whitney U test, separately for each variable . 

After correcting for multiple comparisons, we found no variable to differ between low and high caste dominance villages. There is an effect in the area variable when no multiple comparison correction is applied (p = 0.02), but this effect is gone after applying FDR correction (p = 0.70). Note that the variable is entered as a control variable in the household level analysis reported below as well, so that even if there was an effect here, we still control for area in the main analysis. 

We can conclude that both types of villages only differ in caste dominance, and not in any other variable tested here.

## Analysis 2: Household level

The main analysis to test the claim was performed on household level outcomes. We predict that household income is predicted by village level caste dominance for low caste households. 

Before we can test this claim, we need to clean up and assess the distribution of the main dependent variable: total household income. Only low caste households with no missing data are included in this analysis. 

Please note that since the additional instructions prohibit the use of control variables, the preprocessing step (focused on transforming control variables) reported in Task 1 was omitted here.  

### Outlier detection

First, we removed potential outliers from the data. For this purpose, we first plotted a histogram to assess its distribution. 

```{r household_outcomes_outliers_1}

# to avoid issues with missing and unbalanced data, analyses will be restricted to complete data sets
hh = hh[complete.cases(hh),]

# before detecting outliers, we check the data distribution
hist(hh$totinc, breaks = 100)

```

Total income clearly is not distributed normally, the distribution is highly skewed and has a very long tail. In order to still detect outliers and circumvent the non-normal distribution issue, we log-transformed the income variable, after which the distribution approximated a normal distribution. Then, we applied the 1.5*IQR criterion to detect and remove outliers. 16 outliers were removed this way, leading to a total sample size of n = 943 households. This is identical to the analysis reported in Task 1. 

```{r household_outcomes_outliers_2}

# -> data are not normally distributed, making outlier detection difficult, so we log transform the data
hist(log(hh$totinc), breaks=100)
# -> now that we have an approximate normal distribution, we can detect outliers using the 1.5*IQR criterion
#boxplot(log(hh$totinc))
out = boxplot.stats(log(hh$totinc))$out
out_ind = which(log(hh$totinc) %in% c(out))
#out_ind
hhfilt = hh[-out_ind,]
```

### Data distribution 

In order to test whether total income was distributed normally after outlier removal, we generated a q-q-plot, showing the data as black dots and the fitted normal distribution as a blue line (with the 95% confidence interval shown as a blue area).

```{r household_outcomes_distribution_1}
qqp(hhfilt$totinc,'norm')
```

The normal distribution still fits the data badly, and we tested whether a log-normal distribution would approximate the income data better than the normal distribution did (again using q-q plots)

```{r household_outcomes_distribution_2}
qqp(hhfilt$totinc,'lnorm')
```

Clearly, the log-normal distribution approximates the data much better, and we will use it in the estimation of the multilevel mixed models described below. 

### Multilevel mixed models

In order to directly test the claim, we used multilevel mixed models. The household level data has a nested structure, with households being nested in villages, and villages being nested in states. Multilevel mixed models allow us to describe this nested structure as nested random effects. In contrast to the Task 1 analysis, we only use a 2-level nested structure here, since the use of district controls (a 3rd level nested between villages and states) is prohibited here. 

In order to account for the non-normal distribution of the total income variable, we used a generalized linear mixed model using the log-normal distribution. The model was fitted to the data using quasi-likelihood estimation. 

The fitted model used two random effects (village, state) and one fixed effect (village level caste dominance -> domlow). This tested the H0 = "Village caste dominance does not affect total household income for low caste households."

```{r household_outcomes_mixed_models_1}

# A good introduction on GLMM and PQL:
# https://www.juliapilowsky.com/2018/10/19/a-practical-guide-to-mixed-models-in-r/

# base model 
PQL <- glmmPQL(totinc ~ domlow, 
               ~1 | bihar/village, 
               family = gaussian(link = "log"), 
               data = hhfilt, 
               verbose = FALSE)
summary(PQL)

```

Results indicated that village dominance had a clear effect on total household income (t(66) = 3.05, p = 0.003). 

The additional instruction prohibited the use of the following control variables: district controls, crop controls, distance controls, groundwater controls, public goods controls, irrigation, credit, tenancy, water buyer status. 

However, in the Task 1 model we also used literacy, total land of the village, and type of caste as key control variables. These variables were not included in the additional instruction in Task 2. Therefore, literacy, total land of the village, and type of caste were added to the model as fixed effects, and we tested whether the main effect of caste dominance would remain significant. Since the number of control variables was relatively low, we also added all interaction terms to the model. This model will be used to test the claim reported above.   

```{r household_outcomes_mixed_models_2}
PQL <- glmmPQL(totinc ~ literate * totland * caste * domlow,
               ~1 | bihar/village, 
               family = gaussian(link = "log"), 
               data = hhfilt, 
               verbose = FALSE)
summary(PQL)
```

Results indicated that there was a significant main effect of literacy, and caste (BAC vs OBC vs SC) on total income, all ps < 0.007. There was no significant main effect of caste dominance (domlow), t(66) = -0.87, p = 0.39. 

# Conclusion

We can therefore conclude that under the additional instructions provided for Task 2, the claim is not supported by the data.  

# Glossary 

Three different lower castes are included in this analysis: 

* BAC = backwards agricultural caste 
* OBC = other backwards caste
* SC = scheduled caste 

Code book:

Variables names have been left unaltered to allow a better comparison with the original publication. 

* illand = price of irrigated land
* nilland = price of non-irrigated land
* hhlless = landless households
* village = village id
* bihar = state (Bihar vs Uttar Pradesh)
* hhelec = electrified households
* numhh = number of households
* domhigh = dominated by high caste?
* domlow = dominated by low caste? (inverse of domhigh)
* handpump = main drinking source handpump
* well = main drinking source well
* pavedroad = paved road
* noflood = Almost no land suffers from floods
* noalkal = almost no land suffers from alkalinity
* nowaterlog  = almost no water suffers from waterlogging
* nosoiler = almost no water suffers from soil erosion
* mpaddy =  main crop is paddy
* mwheat = main crop is wheat
* ppaddy = price paddy
* pwheat = price wheat
* bus3 = distance bus stop
* tele3 = distance telephone
* ps3 = distance police station
* pds3 = distance pds shop
* bank3 = distance bank
* pps3 = distance primary school
* ms3 = distance middle school
* ss3 = distance secondary school
* phc3 = distance PHC
* hosp3 = distance hospital
* area = aread in hectares
* naturalwater = access to natural water
* gwdevelopment = ground water development
* rainfall= rainfall
* gwavailability = ground water availability
* river = access to river
* canal = access to canal
* area = area of public infrastructure
* pmixdominant = public infrastructure dominant
* tolaparea = total public area